---
title: "WMH_Xsec"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
date: "Last compiled on `r format(Sys.time(), '%d %B %Y')`"
author: Tracy Melzer
---

This presents the plots and statistical analyses for the total WMH volume calculated using 4 different segmentation algorithms.

# Libraries
```{r, include=FALSE}
# 
#rm(list = ls())
setwd('/home/tracym/Documents/R/WMH_Xsec/')

library(tidyverse)
library(plyr)
library(dplyr)
library(ggplot2)
library(ggfortify)
library("plotly")
library(emmeans)
library(rcompanion)
library(pander)
#Display 2 digists in the lm outputs 
panderOptions("digits",2)
library(lm.beta)
library(ggeffects)
library(ggpubr)
library(car)
library(lme4)

```


# Read in data
```{r, include=FALSE }

dat.base <- read.csv("dat_base_anon.csv")

#Reorder diagnosis
dat.base <- dat.base %>% 
  mutate(diagnosis = fct_relevel(diagnosis, "Control-N", "PD-N", "PD-MCI", "PDD"))


#Load performance metrics as well
p.metrics = read.csv("performance_measures_anon.csv")

p.metrics <- p.metrics %>% 
  mutate(anon_id = factor(anon_id))
```
 
\newpage
# Performance metrics
## Table
Here is a summary table of the performance metrics based on the 40 manually-traced gold standard WMH masks.  

```{r performance metrics, echo=FALSE}

p.metrics %>% 
  group_by(algorithm) %>% 
  summarise(mean_SI = mean(SI), sd_SI = sd(SI), 
            mean_FPRc = mean(FPR_clusters), 
            mean_FNRc = mean(FNR_clusters)) %>% 
  knitr::kable()

```


## Plot
A plot of the Dice Similarity Coefficient across the 4 methods in the training sample of n=40.    
```{r, echo=FALSE, warning=FALSE}

#colours <- c("#6495ED","#00008B","#339900",  "#FF9326", "#D92121")
p.metrics %>% 
  ggplot(aes(x=reorder(algorithm,-SI),y=SI)) +
  geom_boxplot(outlier.shape=NA, width=0.8) +
  #geom_line(aes(group=ID),alpha=0.2)+
  geom_point(aes(color=algorithm), position = position_jitter(width=0.15),size=2.5) +
  ylab("Dice Similarity Coefficient") + xlab("") +
  theme(legend.position = "none") +
  theme(axis.text.x=element_text(face="bold",size=15, color="black")) +
  theme(axis.text.y=element_text(face="bold",size=15, color="black")) +
  theme(axis.title.y=element_text(face="bold",size=15, color="black")) 
ggsave("SI_metrics.pdf",width=8,height=5)


```


## Stats
```{r, echo=FALSE}

#A simple, non-repeated measures ANOVA
#si.mod <- lm(SI ~ algorithm, data=p.metrics)
#pander(si.mod)

#fpr.mod <- lm(FPR ~ algorithm, data=p.metrics)
#pander(fpr.mod)

#fnr.mod <- lm(FNR ~ algorithm, data=p.metrics)
#pander(fnr.mod)

#der.mod <- lm(DER ~ algorithm, data=p.metrics)
#pander(der.mod)



#Run a repeated measured mixed model 
si.mod.rep <- lmer(SI ~ algorithm + (1|anon_id), data=p.metrics)
summary(si.mod.rep)
si.mod.rep.emm <- emmeans(si.mod.rep, tukey ~ algorithm)
plot(si.mod.rep.emm$contrasts,xlab="delta SI")
eff.si <- eff_size(si.mod.rep.emm$contrasts, sigma = sigma(si.mod.rep), edf = df.residual(si.mod.rep), method="identity")
ggsave("SI_pairwise.pdf",width=8,height=5)

cat("Means and CIs for SI")
knitr::kable(si.mod.rep.emm$emmeans)
cat("BIANCA had larger SI relative to the other three methods.\n There was no difference among the other three algorithms.
")
cat("Contrasts")
knitr::kable(si.mod.rep.emm$contrasts)


#FPR
fpr.mod.rep <- lmer(FPR ~ algorithm + (1|anon_id), data=p.metrics)
summary(fpr.mod.rep)
fpr.mod.rep.emm <- emmeans(fpr.mod.rep, tukey ~ algorithm)
plot(fpr.mod.rep.emm$contrasts,xlab="delta SI")
eff.fpr <- eff_size(fpr.mod.rep.emm$contrasts, sigma = sigma(fpr.mod.rep), edf = df.residual(fpr.mod.rep), method="identity")
ggsave("FPR_pairwise.pdf",width=8,height=5)

cat("Means and CIs for FPR")
knitr::kable(fpr.mod.rep.emm$emmeans)


#FNR
fnr.mod.rep <- lmer(FNR ~ algorithm + (1|anon_id), data=p.metrics)
summary(fnr.mod.rep)
fnr.mod.rep.emm <- emmeans(fnr.mod.rep, tukey ~ algorithm)
plot(fnr.mod.rep.emm$contrasts,xlab="delta SI")
eff.fnr <- eff_size(fnr.mod.rep.emm$contrasts, sigma = sigma(fnr.mod.rep), edf = df.residual(fnr.mod.rep), method="identity")
ggsave("FNR_pairwise.pdf",width=8,height=5)
cat("Means and CIs for FNR")
knitr::kable(fnr.mod.rep.emm$emmeans)

#der
der.mod.rep <- lmer(DER ~ algorithm + (1|anon_id), data=p.metrics)
summary(der.mod.rep)
der.mod.rep.emm <- emmeans(der.mod.rep, tukey ~ algorithm)
plot(der.mod.rep.emm$contrasts,xlab="delta SI")
eff.der <- eff_size(der.mod.rep.emm$contrasts, sigma = sigma(der.mod.rep), edf = df.residual(der.mod.rep), method="identity")
ggsave("DER_pairwise.pdf",width=8,height=5)
cat("Means and CIs for DER")
knitr::kable(der.mod.rep.emm$emmeans)



```


\newpage
# Summarise
A few summary demographics cohort.   
## Age
```{r, echo=FALSE}

dat.base %>% 
  group_by(diagnosis) %>% 
  dplyr::summarise(n(), age_m = mean(age), std = sd(age), ed_m = mean(education),ed_std = sd(education) ) %>% 
  knitr::kable()

dat.base %>% 
  group_by(diagnosis) %>%
  filter(global_z!='NA') %>% 
  dplyr::summarise(n(),cog_m = mean(global_z), cog_std = sd(global_z) ) %>% 
  knitr::kable()

dat.base %>% 
  group_by(diagnosis) %>%
  filter(MoCA!='NA') %>% 
  dplyr::summarise(n(),moca_m = mean(MoCA),moca_std = sd(MoCA)) %>% 
  knitr::kable()

dat.base %>% 
  group_by(diagnosis) %>%
  filter(Part_III!='NA') %>% 
  dplyr::summarise(n(),updrs_m = mean(Part_III),updrs_std = sd(Part_III)) %>% 
  knitr::kable()

dat.base %>% 
  group_by(diagnosis) %>%
  filter(H_Y!='NA') %>% 
  dplyr::summarise(n(),hy_m = median(H_Y),hy_min = min(H_Y), hy_max = max(H_Y)) %>% 
  knitr::kable()


```


## Sex  
```{r, echo=FALSE}

dat.base %>% 
  group_by(diagnosis,sex) %>% 
  dplyr::summarise(n()) %>% 
  knitr::kable()

```


# Method Comparison:Baseline
## Data wrangle
```{r, include=FALSE}
dat.plot.base <- dat.base %>% 
  gather(method,wmh, c(`bianca`,`lga`,`lpa`, `ubo`))%>%
  mutate(method=factor(method)) %>% 
  mutate(method = fct_relevel(method,"bianca","lpa", "ubo","lga")) %>% 
  mutate(method = fct_recode(method,BIANCA="bianca",LPA="lpa", UBO="ubo",LGA="lga")) %>% 
  mutate(diagnosis = fct_relevel(diagnosis, "Control-N", "PD-N", "PD-MCI", "PDD"))%>% 
  mutate(diagnosis = revalue(diagnosis,c("Control-N" = "Control"))) %>% 
  mutate(wmh.icv = wmh/ICV) 

```


\newpage
## Figure: PD vs HC
Plot PD vs controls for the 4 methods.  
\hfill\break
```{r, echo=FALSE, warning=FALSE}

colours <- c("#6495ED", "#339900",  "#FF9326","#D92121")
dat.plot.base %>% 
  ggplot(aes(x=np_group,y=(wmh+1))) +
  #geom_violin()+
  geom_boxplot(outlier.shape=NA, width=0.8) +
  geom_point(aes(color=diagnosis), position = position_jitter(width=0.15), alpha=0.6, size=1) +
  scale_colour_manual (values=colours,name = "diagnosis") + 
  xlab("") +
  scale_y_log10('1 + White Matter Hyperintensities (ml)',limits=c(0.9, 100)) +
  facet_wrap(~method,nrow = 1 ) +
  theme(axis.text.x = element_text(angle=-45,hjust=0.1))
  ggsave("WMH_Group_4method_facet_method.pdf",width=10,height=5)
  
```

\newpage
## Figure: ADJUSTED PD vs HC
Plot the residuals from a model adjusting for age, sex, and ICV.  
\hfill\break
```{r, echo=FALSE, warning=FALSE}

#Copied from Figure: ADJUSTED Diagnosis, as it is the same modle (without diagnosis or group)
#Fit a model without Group, then use those residuals in the plot.
diag.lga <- lm(log(lga+1)~age_s10 + sex + icv_dm ,dat=dat.base)
diag.lpa <- lm(log(lpa+1)~age_s10 + sex + icv_dm ,dat=dat.base)
diag.bianca <- lm(log(bianca+1)~age_s10 + sex + icv_dm ,dat=dat.base)
diag.ubo <- lm(log(ubo+1)~age_s10 + sex + icv_dm ,dat=dat.base)

#Add the intercept back in to all residuals
dat.base$resid.diag.lga <- residuals(diag.lga) + diag.lga$coefficients[1]
dat.base$resid.diag.lpa <- residuals(diag.lpa) + diag.lpa$coefficients[1]
dat.base$resid.diag.bianca <- residuals(diag.bianca) + diag.bianca$coefficients[1]
dat.base$resid.diag.ubo <- residuals(diag.ubo) + diag.ubo$coefficients[1]

dat.plot.base.pd.adjust <- dat.base %>% 
  #gather(method,wmh, c(`bianca`,`lga`,`lpa`, `ubo`) )%>%
  gather(method,resid, c(`resid.diag.bianca`,`resid.diag.lga`,`resid.diag.lpa`, `resid.diag.ubo`))%>%
  mutate(method=factor(method)) %>% 
  mutate(method = factor(method, labels = c('BIANCA', 'LGA', 'LPA', 'UBO'))) %>% 
  mutate(method = fct_relevel(method,"BIANCA","LPA", "UBO","LGA")) %>% 
  mutate(diagnosis = fct_relevel(diagnosis, "Control-N", "PD-N", "PD-MCI", "PDD"))%>% 
  mutate(diagnosis = revalue(diagnosis,c("Control-N" = "Control"))) 
 
  
colours <- c("#6495ED", "#339900",  "#FF9326","#D92121")
dat.plot.base %>%
  ggplot(aes(x=np_group,y=wmh+1)) +
  #geom_violin(data = dat.plot.base.pd.adjust, aes(x=np_group,y=exp(resid)), fill='gray90', alpha=0.8, trim=FALSE)+
  geom_boxplot(data = dat.plot.base.pd.adjust, aes(x=np_group,y=exp(resid)), alpha=0.8, outlier.shape=NA,
               width=0.5,lwd=0.3)+
  geom_point(aes(color=diagnosis), position = position_jitter(width=0.1), alpha=0.4, size=1) +
  scale_colour_manual (values=colours,name = "") +
  xlab("") +
  scale_y_log10('1+White Matter Hyperintensities (ml)',limits=c(0.9, 100)) +
  facet_wrap(~method,nrow=1)+
  theme(legend.position="none")
  ggsave("WMH_adjusted_box_GROUP_LINEAR_4method_facet_method.pdf",width=8,height=5)

  
```



\newpage
## Figure: Diagnosis by Method
\hfill\break
```{r, echo=FALSE, warning=FALSE}

#colours <- c("#6495ED","#00008B","#339900",  "#FF9326","#D92121")
colours <- c("#6495ED", "#339900",  "#FF9326","#D92121")
dat.plot.base %>% 
  ggplot(aes(x=diagnosis,y=(wmh+1))) +
  #geom_violin()+
  geom_boxplot(outlier.shape=NA, width=0.8) +
  geom_point(aes(color=diagnosis), position = position_jitter(width=0.15), alpha=0.6, size=1) +
  scale_colour_manual (values=colours,name = "diagnosis") + 
  xlab("") +
  scale_y_log10('1 + White Matter Hyperintensities (ml)',limits=c(0.9, 100)) +
  facet_wrap(~method) +
  theme(axis.text.x = element_text(angle=-45,hjust=0.1), legend.title=element_blank())
  ggsave("WMH_4method_facet_method.pdf",width=10,height=5)

  
```

\newpage
## Figure: ADJUSTED Diagnosis by Method  
Plot the residuals from a model adjusting for age, sex, and ICV.  
\hfill\break
```{r, echo=FALSE, warning=FALSE}

#Same as above, but repeat here in case the other hasn't been run.
#Fit a model without Diagnosis, then use those residuals in the plot.
diag.lga <- lm(log(lga+1)~age_s10 + sex + icv_dm ,dat=dat.base)
diag.lpa <- lm(log(lpa+1)~age_s10 + sex + icv_dm ,dat=dat.base)
diag.bianca <- lm(log(bianca+1)~age_s10 + sex + icv_dm ,dat=dat.base)
diag.ubo <- lm(log(ubo+1)~age_s10 + sex + icv_dm ,dat=dat.base)

dat.base$resid.diag.lga <- residuals(diag.lga) + diag.lga$coefficients[1]
dat.base$resid.diag.lpa <- residuals(diag.lpa) + diag.lpa$coefficients[1]
dat.base$resid.diag.bianca <- residuals(diag.bianca) + diag.bianca$coefficients[1]
dat.base$resid.diag.ubo <- residuals(diag.ubo) + diag.ubo$coefficients[1]

dat.plot.base.adjust <- dat.base %>% 
  #gather(method,wmh, c(`bianca`,`lga`,`lpa`, `ubo`) )%>%
  gather(method,resid, c(`resid.diag.bianca`,`resid.diag.lga`,`resid.diag.lpa`, `resid.diag.ubo`))%>%
  mutate(method=factor(method)) %>% 
  mutate(method = factor(method, labels = c('BIANCA', 'LGA', 'LPA', 'UBO'))) %>% 
  mutate(method = fct_relevel(method,"BIANCA","LPA", "UBO","LGA")) %>% 
  mutate(diagnosis = fct_relevel(diagnosis, "Control-N", "PD-N", "PD-MCI", "PDD"))%>% 
  mutate(diagnosis = revalue(diagnosis,c("Control-N" = "Control"))) 
 

  dat.plot.base %>%
  ggplot(aes(x=diagnosis,y=wmh+1)) +
  #geom_violin(data = dat.plot.base.adjust, aes(x=diagnosis,y=exp(resid)), fill='grey90', alpha=0.8, trim=FALSE)+
  geom_boxplot(data = dat.plot.base.adjust, aes(x=diagnosis,y=exp(resid)), alpha=0.8, outlier.shape=NA, width=0.5,lwd=0.3)+
  geom_point(aes(color=diagnosis), position = position_jitter(width=0.1), alpha=0.4, size=1) +
  scale_colour_manual (values=colours,name = "") +
  xlab("") +
  #ylab("log(1+White Matter Hyperintensities)") +
  scale_y_log10('1+White Matter Hyperintensities (ml)',limits=c(0.9, 100)) +
  facet_wrap(~method,nrow=1) +
  theme(axis.text.x = element_text(angle=-45,hjust=0.1), legend.position="none")
  ggsave("WMH_adjusted_box_DIAGNOSIS_LINEAR_4method_facet_method.pdf",width=8,height=5)

```


\hfill\break
\newpage
# ANALYSIS
## Group
Simple PD vs Control comparisons across the 4 WMH segmentation algorithms.  
```{r, echo=FALSE }

lm.pd.lga <- lm(log(lga+1) ~ age_s10 + sex + icv_dm + group, data=dat.base)
#summary(lm.pd.lga)
#pander(lm.pd.lga)

#Use the lm.beta package to produce standardized betas to allow a better comparison across the different models.
lm.pd.lga.beta <- lm.beta(lm.pd.lga)
pander(lm.pd.lga.beta)

#Effect size
lm.pd.lga.emm <- emmeans(lm.pd.lga, tukey ~ group)
eff.pd.lga <- eff_size(lm.pd.lga.emm$contrasts,sigma = sigma(lm.pd.lga), edf = df.residual(lm.pd.lga), method="identity")


lm.pd.lpa <- lm(log(lpa+1) ~ age_s10 + sex + icv_dm + group, data=dat.base)
#summary(lm.pd.lpa)
#pander(lm.pd.lpa)
lm.pd.lpa.beta <- lm.beta(lm.pd.lpa)
pander(lm.pd.lpa.beta)

#Effect size
lm.pd.lpa.emm <- emmeans(lm.pd.lpa, tukey ~ group)
eff.pd.lpa <- eff_size(lm.pd.lpa.emm$contrasts,sigma = sigma(lm.pd.lpa), edf = df.residual(lm.pd.lpa), method="identity")



lm.pd.bianca <- lm(log(bianca+1) ~ age_s10 + sex + icv_dm + group, data=dat.base)
#summary(lm.pd.bianca)
#pander(lm.pd.bianca)
lm.pd.bianca.beta <- lm.beta(lm.pd.bianca)
pander(lm.pd.bianca.beta)

#Effect size
lm.pd.bianca.emm <- emmeans(lm.pd.bianca, tukey ~ group)
eff.pd.bianca <- eff_size(lm.pd.bianca.emm$contrasts,sigma = sigma(lm.pd.bianca), edf = df.residual(lm.pd.bianca), method="identity")




lm.pd.ubo <- lm(log(ubo+1) ~ age_s10 + sex + icv_dm + group, data=dat.base)
#summary(lm.pd.ubo)
#pander(lm.pd.ubo)
lm.pd.ubo.beta <- lm.beta(lm.pd.ubo)
pander(lm.pd.ubo.beta)

#Effect size
lm.pd.ubo.emm <- emmeans(lm.pd.ubo, tukey ~ group)
eff.pd.ubo <- eff_size(lm.pd.ubo.emm$contrasts,sigma = sigma(lm.pd.ubo), edf = df.residual(lm.pd.ubo), method="identity")





cat("Standardized Betas for each method")
#Display the standardized betas for each method
print(lm.pd.lpa.beta)
print(lm.pd.lga.beta)
print(lm.pd.bianca.beta)
print(lm.pd.ubo.beta)



```

\newpage
## Diagnosis
```{r, echo=FALSE, results='asis', fig.width=4,fig.height=3 }

##LGA
##**********************
lm.diagnosis.lga <- lm(log(lga+1) ~ age_s10 + sex + icv_dm + diagnosis, data=dat.base)
#summary(lm.diagnosis.lga)
#pander(lm.diagnosis.lga)


#Get standardized betas
lm.diagnosis.lga <- lm.beta(lm.diagnosis.lga)
pander(lm.diagnosis.lga)

#cat("\n\n\\newpage\n")
lm.diagnosis.lga.pairwise <- ref_grid(lm.diagnosis.lga)
lm.diagnosis.lga.emm <- emmeans(lm.diagnosis.lga.pairwise, tukey ~ diagnosis)

#Effect size
eff.diag.lga <- eff_size(lm.diagnosis.lga.emm$contrasts,sigma = sigma(lm.diagnosis.lga), edf = df.residual(lm.diagnosis.lga), method="identity")

plot(lm.diagnosis.lga.emm, xlab="LGA")
ggsave("LGA_emm.pdf",width=8,height=5)

cat(" \n")
plot(lm.diagnosis.lga.emm$contrasts,xlab="LGA")
ggsave("LGA_diagnosis_pairwise.pdf",width=8,height=5)

##******************************************************


cat("\n\n\\newpage\n")


##LPA
##*********************************************************
lm.diagnosis.lpa <- lm(log(lpa+1) ~ age_s10 + sex + icv_dm + diagnosis, data=dat.base)
#pander(lm.diagnosis.lpa)

#Get standardized betas
lm.diagnosis.lpa <- lm.beta(lm.diagnosis.lpa)
pander(lm.diagnosis.lpa)

lm.diagnosis.lpa.pairwise <- ref_grid(lm.diagnosis.lpa)
lm.diagnosis.lpa.emm <- emmeans(lm.diagnosis.lpa.pairwise, tukey ~ diagnosis)

#Effect size
eff.diag.lpa <- eff_size(lm.diagnosis.lpa.emm$contrasts,sigma = sigma(lm.diagnosis.lpa), edf = df.residual(lm.diagnosis.lpa), method="identity")

plot(lm.diagnosis.lpa.emm,xlab="LPA")
ggsave("LPA_emm.pdf",width=8,height=5)

cat(" \n")
plot(lm.diagnosis.lpa.emm$contrasts,xlab="LPA")
ggsave("LPA_diagnosis_pairwise.pdf",width=8,height=5)

##******************************************************


cat("\n\n\\newpage\n")


##BIANCA
##**************************************************
lm.diagnosis.bianca <- lm(log(bianca+1) ~ age_s10 + sex + icv_dm + diagnosis, data=dat.base)
#pander(lm.diagnosis.bianca)


#Get standardized betas
lm.diagnosis.bianca <- lm.beta(lm.diagnosis.bianca)
pander(lm.diagnosis.bianca)

lm.diagnosis.bianca.pairwise <- ref_grid(lm.diagnosis.bianca)
lm.diagnosis.bianca.emm <- emmeans(lm.diagnosis.bianca.pairwise, tukey ~ diagnosis)

eff.diag.bianca <- eff_size(lm.diagnosis.bianca.emm$contrasts,sigma = sigma(lm.diagnosis.bianca), edf = df.residual(lm.diagnosis.bianca), method="identity")

plot(lm.diagnosis.bianca.emm, xlab="BIANCA")
ggsave("BIANCA_emm.pdf",width=8,height=5)

#Adds a line break in the pdf
cat(" \n")
plot(lm.diagnosis.bianca.emm$contrasts,  xlab="BIANCA")
ggsave("BIANCA_diagnosis_pairwise.pdf",width=8,height=5)

##******************************************************

cat("\n\n\\newpage\n")

##UBO
##*************************************************
lm.diagnosis.ubo <- lm(log(ubo+1) ~ age_s10 + sex + icv_dm + diagnosis, data=dat.base)
#pander(lm.diagnosis.ubo)


lm.diagnosis.ubo <- lm.beta(lm.diagnosis.ubo)
pander(lm.diagnosis.ubo)

lm.diagnosis.ubo.pairwise <- ref_grid(lm.diagnosis.ubo)
lm.diagnosis.ubo.emm <- emmeans(lm.diagnosis.ubo.pairwise, tukey ~ diagnosis)

#Effect size
eff.diag.ubo <- eff_size(lm.diagnosis.ubo.emm$contrasts,sigma = sigma(lm.diagnosis.ubo), edf = df.residual(lm.diagnosis.ubo), method="identity")

plot(lm.diagnosis.ubo.emm, xlab="UBO")
ggsave("UBO_emm.pdf",width=8,height=5)

#Adds a line in the PDF
cat(" \n")
plot(lm.diagnosis.ubo.emm$contrasts, xlab="UBO")
ggsave("UBO_diagnosis_pairwise.pdf",width=8,height=5)

##******************************************************



```


\newpage
## Diagnosis pairwise comp
\hfill\break
Pairwise comparisons
```{r echo=FALSE, results='asis', fig.width=6,fig.height=5}


#Try for a single plot, facetted by method

lga.diag.dat <- as.data.frame(confint(lm.diagnosis.lga.emm$contrasts)) %>% 
  mutate(contrast = fct_relevel(contrast, "Control-N - PD-N", "Control-N - PD-MCI", "Control-N - PDD", "PD-N - PD-MCI", "PD-N - PDD", "PD-MCI - PDD")) %>% 
  mutate(method = 'lga')

lpa.diag.dat <- as.data.frame(confint(lm.diagnosis.lpa.emm$contrasts)) %>% 
  mutate(contrast = fct_relevel(contrast, "Control-N - PD-N", "Control-N - PD-MCI", "Control-N - PDD", "PD-N - PD-MCI", "PD-N - PDD", "PD-MCI - PDD")) %>% 
  mutate(method = 'lpa')

bianca.diag.dat <- as.data.frame(confint(lm.diagnosis.bianca.emm$contrasts)) %>% 
  mutate(contrast = fct_relevel(contrast, "Control-N - PD-N", "Control-N - PD-MCI", "Control-N - PDD", "PD-N - PD-MCI", "PD-N - PDD", "PD-MCI - PDD")) %>% 
  mutate(method = 'bianca')

ubo.diag.dat <- as.data.frame(confint(lm.diagnosis.ubo.emm$contrasts)) %>% 
  mutate(contrast = fct_relevel(contrast, "Control-N - PD-N", "Control-N - PD-MCI", "Control-N - PDD", "PD-N - PD-MCI", "PD-N - PDD", "PD-MCI - PDD")) %>% 
  mutate(method = 'ubo')

####***************************************


#Bind together
diag.mod.plot <- rbind(lga.diag.dat, lpa.diag.dat, bianca.diag.dat, ubo.diag.dat)
diag.mod.plot <- diag.mod.plot %>% 
  mutate(method = fct_relevel(method, "lga", "lpa", "bianca", "ubo"))%>% 
  mutate(method = factor(method, labels = c('LGA', 'LPA', 'BIANCA', 'UBO'))) %>%
  mutate(method = fct_relevel(method,"BIANCA","LPA", "UBO","LGA")) %>% 
  mutate(contrast = fct_recode(contrast, "Control - PD-N" = 'Control-N - PD-N', "Control - PD-MCI" = "Control-N - PD-MCI", "Control - PDD"= "Control-N - PDD" ))


#All together
diag.mod.plot %>% 
    ggplot(aes(x=contrast, y=estimate)) + 
  geom_crossbar(aes(ymax=upper.CL, ymin=lower.CL), fill='olivedrab4', alpha=0.5, fatten=0, width=0.2, linetype=0) +
    geom_point(size=2.5)+
  coord_flip() +
  geom_hline(yintercept = 0)+
  xlab("")+
  ylab("Estimate") +
  facet_wrap(~method) 
ggsave("WMH_DIAG_Tukey_4method_facet_method.pdf",width=8,height=5)

```


\newpage
## Cogz in PD
\hfill\break
Regressions between log(WMH+1) and Cogz aross the 4 methods.
There is a model comparison within-method to determine the best model. The model with cogz~log(wmh+1) turns out to be the best (with the smallest AIC and BIC) in all 4 cases.
```{r, echo=FALSE}

#Run just in PD
dat.base.pd <- dat.base %>% 
  filter(group=='PD') %>% 
  mutate(group=factor(group)) %>% 
  mutate(diagnosis=factor(diagnosis))

#2 step model comparison. 1st, confirm which of the models within each method is the best, then compare the best model across methods to each other. 

#***LGA***
########################################
cogz.pd.lga.r <- lm(global_z ~ age_s10 + sex + lga.icv, data=dat.base.pd )
pander(cogz.pd.lga.r)

cogz.pd.lga.raw <- lm(global_z ~ age_s10 + sex + lga, data=dat.base.pd )
pander(cogz.pd.lga.raw)


cogz.pd.lga.ln <- lm(global_z ~ age_s10 + sex + log(lga+1), data=dat.base.pd )
pander(cogz.pd.lga.ln)

cogz.pd.lga.ln.icv <- lm(global_z ~ age_s10 + sex + log(lga+1) + icv_dm, data=dat.base.pd )
pander(cogz.pd.lga.ln.icv)

#Compare across the three within-method models
anova(cogz.pd.lga.r, cogz.pd.lga.raw, cogz.pd.lga.ln, cogz.pd.lga.ln.icv)

#Get AIC for within-method models
pander(compareLM(cogz.pd.lga.r, cogz.pd.lga.raw, cogz.pd.lga.ln, cogz.pd.lga.ln.icv))

########################################


#***LPA***
########################################
cogz.pd.lpa.r <- lm(global_z ~ age_s10 + sex + lpa.icv, data=dat.base.pd )
pander(cogz.pd.lpa.r)

cogz.pd.lpa.ln <- lm(global_z ~ age_s10 + sex + log(lpa+1), data=dat.base.pd )
pander(cogz.pd.lpa.ln)

cogz.pd.lpa.ln.icv <- lm(global_z ~ age_s10 + sex + log(lpa+1) + icv_dm, data=dat.base.pd )
pander(cogz.pd.lpa.ln.icv)

lpa.pd.cogz <- lm(log(lpa+1) ~ age_s10 + sex + global_z + icv_dm, data=dat.base.pd )
pander(lpa.pd.cogz)


#Compare across the three within-method models
anova(cogz.pd.lpa.r, cogz.pd.lpa.ln, cogz.pd.lpa.ln.icv)

#Get AIC for within-method models
pander(compareLM(cogz.pd.lpa.r, cogz.pd.lpa.ln, cogz.pd.lpa.ln.icv))
########################################


#***BIANCA***
########################################
cogz.pd.bianca.r <- lm(global_z ~ age_s10 + sex + bianca.icv, data=dat.base.pd )
pander(cogz.pd.bianca.r)

cogz.pd.bianca.ln <- lm(global_z ~ age_s10 + sex + log(bianca+1), data=dat.base.pd )
pander(cogz.pd.bianca.ln)

cogz.pd.bianca.ln.icv <- lm(global_z ~ age_s10 + sex + log(bianca+1) + icv_dm, data=dat.base.pd )
pander(cogz.pd.bianca.ln.icv)

#Compare across the three within-method models
anova(cogz.pd.bianca.r, cogz.pd.bianca.ln, cogz.pd.bianca.ln.icv)

#Get AIC for within-method models
pander(compareLM(cogz.pd.bianca.r, cogz.pd.bianca.ln, cogz.pd.bianca.ln.icv))

########################################


#***UBO***
########################################
cogz.pd.ubo.r <- lm(global_z ~ age_s10 + sex + ubo.icv, data=dat.base.pd )
pander(cogz.pd.ubo.r)

cogz.pd.ubo.ln <- lm(global_z ~ age_s10 + sex + log(ubo+1), data=dat.base.pd )
pander(cogz.pd.ubo.ln)

cogz.pd.ubo.ln.icv <- lm(global_z ~ age_s10 + sex + log(ubo+1) + icv_dm, data=dat.base.pd )
pander(cogz.pd.ubo.ln.icv)

#Compare across the three within-method models
anova(cogz.pd.ubo.r, cogz.pd.ubo.ln, cogz.pd.ubo.ln.icv)

#Get AIC for within-method models
pander(compareLM(cogz.pd.ubo.r, cogz.pd.ubo.ln, cogz.pd.ubo.ln.icv))

########################################


  
```


\newpage
### Cogz standardized betas  
This calculates and reports the standardized betas for the model with the lowest AIC for each method.  
```{r, echo=FALSE}

cogz.pd.lga.ln <- lm.beta(cogz.pd.lga.ln)
pander(cogz.pd.lga.ln)
print(cogz.pd.lga.ln)

cogz.pd.lpa.ln <- lm.beta(cogz.pd.lpa.ln)
pander(cogz.pd.lpa.ln)
print(cogz.pd.lpa.ln)

cogz.pd.bianca.ln <- lm.beta(cogz.pd.bianca.ln)
pander(cogz.pd.bianca.ln)
print(cogz.pd.bianca.ln)

cogz.pd.ubo.ln <- lm.beta(cogz.pd.ubo.ln)
pander(cogz.pd.ubo.ln)
print(cogz.pd.ubo.ln)


```

\newpage
## UPDRS in PD
\hfill\break
Regressions between log(WMH+1) and UPDRS aross the 4 methods.
There is a model comparison within-method to determine the best model. The model with UPDRS~log(wmh+1) turns out to be the best (with the smallest AIC and BIC) in all 4 cases.
```{r, echo=FALSE}

#Run just in PD
dat.base.pd <- dat.base %>% 
  filter(group=='PD') %>% 
  mutate(group=factor(group)) %>% 
  mutate(diagnosis=factor(diagnosis))

#2 step model comparison. 1st, confirm which of the models within each method is the best, then compare the best model across methods to each other. 

#***LGA***
########################################
updrs.pd.lga.r <- lm(Part_III ~ age_s10 + sex + lga.icv, data=dat.base.pd )
pander(updrs.pd.lga.r)

updrs.pd.lga.raw <- lm(Part_III ~ age_s10 + sex + lga, data=dat.base.pd )
pander(updrs.pd.lga.raw)

updrs.pd.lga.both <- lm(Part_III ~ age_s10 + sex + lga + log(lga+1), data=dat.base.pd )
pander(updrs.pd.lga.both)

updrs.pd.lga.ln <- lm(Part_III ~ age_s10 + sex + log(lga+1), data=dat.base.pd )
pander(updrs.pd.lga.ln)

updrs.pd.lga.ln.icv <- lm(Part_III ~ age_s10 + sex + log(lga+1) + icv_dm, data=dat.base.pd )
pander(updrs.pd.lga.ln.icv)

#Compare across the three within-method models
anova(updrs.pd.lga.r, updrs.pd.lga.raw, updrs.pd.lga.ln, updrs.pd.lga.ln.icv)

#Get AIC for within-method models
pander(compareLM(updrs.pd.lga.r, updrs.pd.lga.raw, updrs.pd.lga.both, updrs.pd.lga.ln, updrs.pd.lga.ln.icv))

########################################


#***LPA***
########################################
updrs.pd.lpa.r <- lm(Part_III ~ age_s10 + sex + lpa.icv, data=dat.base.pd )
pander(updrs.pd.lpa.r)

updrs.pd.lpa.ln <- lm(Part_III ~ age_s10 + sex + log(lpa+1), data=dat.base.pd )
pander(updrs.pd.lpa.ln)

updrs.pd.lpa.ln.icv <- lm(Part_III ~ age_s10 + sex + log(lpa+1) + icv_dm, data=dat.base.pd )
pander(updrs.pd.lpa.ln.icv)

lpa.pd.updrs <- lm(log(lpa+1) ~ age_s10 + sex + Part_III + icv_dm, data=dat.base.pd )
pander(lpa.pd.updrs)

#Compare across the three within-method models
anova(updrs.pd.lpa.r, updrs.pd.lpa.ln, updrs.pd.lpa.ln.icv)

#Get AIC for within-method models
pander(compareLM(updrs.pd.lpa.r, updrs.pd.lpa.ln, updrs.pd.lpa.ln.icv))

########################################


#***BIANCA***
########################################
updrs.pd.bianca.r <- lm(Part_III ~ age_s10 + sex + bianca.icv, data=dat.base.pd )
pander(updrs.pd.bianca.r)

updrs.pd.bianca.ln <- lm(Part_III ~ age_s10 + sex + log(bianca+1), data=dat.base.pd )
pander(updrs.pd.bianca.ln)

updrs.pd.bianca.ln.icv <- lm(Part_III ~ age_s10 + sex + log(bianca+1) + icv_dm, data=dat.base.pd )
pander(updrs.pd.bianca.ln.icv)

#Compare across the three within-method models
anova(updrs.pd.bianca.r, updrs.pd.bianca.ln, updrs.pd.bianca.ln.icv)

#Get AIC for within-method models
pander(compareLM(updrs.pd.bianca.r, updrs.pd.bianca.ln, updrs.pd.bianca.ln.icv))

########################################


#***UBO***
########################################
updrs.pd.ubo.r <- lm(Part_III ~ age_s10 + sex + ubo.icv, data=dat.base.pd )
pander(updrs.pd.ubo.r)

updrs.pd.ubo.ln <- lm(Part_III ~ age_s10 + sex + log(ubo+1), data=dat.base.pd )
pander(updrs.pd.ubo.ln)

updrs.pd.ubo.ln.icv <- lm(Part_III ~ age_s10 + sex + log(ubo+1) + icv_dm, data=dat.base.pd )
pander(updrs.pd.ubo.ln.icv)

#Compare across the three within-method models
anova(updrs.pd.ubo.r, updrs.pd.ubo.ln, updrs.pd.ubo.ln.icv)

#Get AIC for within-method models
pander(compareLM(updrs.pd.ubo.r, updrs.pd.ubo.ln, updrs.pd.ubo.ln.icv))

########################################


  
```


\newpage
### UPDRS standardized betas  
This calculates and reports the standardized betas for the model with the lowest AIC for each method.  
```{r, echo=FALSE}

updrs.pd.lga.ln <- lm.beta(updrs.pd.lga.ln)
pander(updrs.pd.lga.ln)
print(updrs.pd.lga.ln)

updrs.pd.lpa.ln <- lm.beta(updrs.pd.lpa.ln)
pander(updrs.pd.lpa.ln)
print(updrs.pd.lpa.ln)

updrs.pd.bianca.ln <- lm.beta(updrs.pd.bianca.ln)
pander(updrs.pd.bianca.ln)
print(updrs.pd.bianca.ln)

updrs.pd.ubo.ln <- lm.beta(updrs.pd.ubo.ln)
pander(updrs.pd.ubo.ln)
print(updrs.pd.ubo.ln)


```

\newpage
# Figures
## Cogz in PD
\hfill\break
```{r, echo=FALSE, warning=FALSE}

#Not adjusted for age or sex (age explains most of the variance.)
colours <- c( "#339900",  "#FF9326","#D92121")
dat.plot.base %>% 
  filter(np_group=='PD') %>% 
  ggplot(aes(x=(wmh+1),y=global_z)) +
  #geom_violin()+
  geom_point(aes(color=diagnosis), alpha=0.6) +
  scale_colour_manual (values=colours,name = "diagnosis") + 
  scale_x_log10('1 + White Matter Hyperintensities (ml)',limits=c(0.9, 100)) +
  ylab("Global cognitive ability") +
  facet_wrap(~method) +
  stat_smooth(method="lm", color='black') 
  #theme(axis.text.x = element_text(angle=-45,hjust=0.1))
  ggsave("Cogz-WMH_4method_facet_method.pdf",width=10,height=5)

  

```

\newpage
### Account for covariates
This plot adjusts for age and sex.
\hfill\break
```{r, echo=FALSE, warning=FALSE}

#Plot Cogz ~ wmh, adjusting for age and sex. 

dat.base.pd.clean <- dat.base.pd %>% 
  filter(global_z!='NA')

z.base.pd <- lm(global_z ~ age_s10 + sex, data=dat.base.pd.clean)
dat.base.pd.clean$resid <- residuals(z.base.pd) + z.base.pd$coefficients[1]

dat.plot.base.clean <- dat.base.pd.clean %>% 
  gather(method,wmh, c(`bianca`,`lga`,`lpa`, `ubo`))%>%
  mutate(method=factor(method)) %>% 
  mutate(method = factor(method, labels = c('BIANCA', 'LGA', 'LPA', 'UBO'))) %>%
  mutate(method = fct_relevel(method,"BIANCA","LPA", "UBO","LGA")) %>%
  mutate(diagnosis = fct_relevel(diagnosis, "PD-N", "PD-MCI", "PDD"))%>% 
  mutate(wmh.icv = wmh/ICV)

  
  colours <- c( "#339900",  "#FF9326","#D92121")
  dat.plot.base.clean %>% 
  filter(np_group=='PD') %>% 
  ggplot(aes(x=(wmh+1),y=global_z)) +
  #geom_violin()+
  geom_point(aes(color=diagnosis), alpha=0.6) +
  scale_colour_manual (values=colours,name = "diagnosis") + 
  scale_x_log10('1 + White Matter Hyperintensities (ml)',limits=c(0.9, 100)) +
  ylab("Global cognitive ability (z score)") +
  facet_wrap(~method,nrow=1) +
  stat_smooth(aes(x=(wmh+1),y=resid),method="lm", color='black') +
  theme(legend.position="none")
  ggsave("Cogz-WMH_ADJUSTED_LINEAR_4method_facet_method.pdf",width=8,height=3.5)
  
 

```

\newpage
## UPDRS in PD
### Account for covariates
This plot adjusts for age and sex.
\hfill\break
```{r, echo=FALSE, warning=FALSE}

#Plot Cogz ~ wmh, adjusting for age and sex. 

dat.base.pd.updrs <- dat.base.pd %>% 
  filter(Part_III!='NA')

updrs.base.pd <- lm(Part_III ~ age_s10 + sex, data=dat.base.pd.updrs)
dat.base.pd.updrs$resid <- residuals(updrs.base.pd) + updrs.base.pd$coefficients[1]

dat.base.pd.updrs <- dat.base.pd.updrs %>% 
  gather(method,wmh, c(`bianca`,`lga`,`lpa`, `ubo`))%>%
  mutate(method=factor(method)) %>% 
  mutate(method = factor(method, labels = c('BIANCA', 'LGA', 'LPA', 'UBO'))) %>%
  mutate(method = fct_relevel(method,"BIANCA","LPA", "UBO","LGA")) %>%
  mutate(diagnosis = fct_relevel(diagnosis, "PD-N", "PD-MCI", "PDD"))%>% 
  mutate(wmh.icv = wmh/ICV)

  
  dat.base.pd.updrs %>% 
  filter(np_group=='PD') %>% 
  ggplot(aes(x=(wmh+1),y=Part_III)) +
  #geom_violin()+
  geom_point(aes(color=diagnosis), alpha=0.6) +
  scale_colour_manual (values=colours,name = "diagnosis") + 
  scale_x_log10('1 + White Matter Hyperintensities (ml)',limits=c(0.9, 100)) +
  ylab("Motor impairment (UPDRS-III)") +
  facet_wrap(~method,nrow=1) +
  stat_smooth(aes(x=(wmh+1),y=resid),method="lm", color='black') +
  theme(legend.position="none")
  ggsave("UPDRS-WMH_ADJUSTED_LINEAR_4method_facet_method.pdf",width=8,height=3.5)

```


# Plots for Paper
## BIANCA plots
```{r, echo=FALSE, warning=FALSE}

#PD vs Control
colours <- c("#6495ED", "#339900",  "#FF9326","#D92121")
dat.plot.base.pd.adjust.bianca <- dat.plot.base.pd.adjust  %>% 
  filter(method=='BIANCA') %>% 
  mutate(method = factor(method))

p1 <- dat.plot.base %>%
  filter(method=='BIANCA') %>%
  ggplot(aes(x=np_group,y=wmh+1)) +
  #geom_violin(data = dat.plot.base.pd.adjust, aes(x=np_group,y=exp(resid)), fill='gray90', alpha=0.8, trim=FALSE)+
  geom_boxplot(data = dat.plot.base.pd.adjust.bianca, aes(x=np_group,y=exp(resid)), alpha=0.8, outlier.shape=NA,
               width=0.5,lwd=0.3)+
  geom_point(aes(color=diagnosis), position = position_jitter(width=0.1), alpha=0.4, size=1) +
  scale_colour_manual (values=colours,name = "") +
  xlab("") +
  scale_y_log10('1+White Matter Hyperintensities (ml)',limits=c(0.9, 100)) +
  facet_wrap(~method,nrow=1)+
  theme(legend.position="none")



#######Diagnosis
colours <- c("#6495ED", "#339900",  "#FF9326","#D92121")

dat.plot.base.adjust.bianca <- dat.plot.base.adjust %>% 
  filter(method=='BIANCA') %>% 
  mutate(method = factor(method)) 
  
p2 <-dat.plot.base %>%
    filter(method=='BIANCA') %>% 
  ggplot(aes(x=diagnosis,y=wmh+1)) +
  #geom_violin(data = dat.plot.base.adjust, aes(x=diagnosis,y=exp(resid)), fill='grey90', alpha=0.8, trim=FALSE)+
  geom_boxplot(data = dat.plot.base.adjust.bianca, aes(x=diagnosis,y=exp(resid)), alpha=0.8, outlier.shape=NA, width=0.5,lwd=0.3)+
  geom_point(aes(color=diagnosis), position = position_jitter(width=0.1), alpha=0.4, size=1) +
  scale_colour_manual (values=colours,name = "") +
  xlab("") +
  #ylab("log(1+White Matter Hyperintensities)") +
  scale_y_log10('1+White Matter Hyperintensities (ml)',limits=c(0.9, 100)) +
  #facet_wrap(~method,nrow=1) +
  theme(axis.text.x = element_text(angle=-45,hjust=0.1), legend.position="none")


##############Pairwise
p2.5 <- diag.mod.plot %>% 
  filter(method=='BIANCA') %>% 
    ggplot(aes(x=contrast, y=estimate)) + 
  geom_crossbar(aes(ymax=upper.CL, ymin=lower.CL), fill='olivedrab4', alpha=0.5, fatten=0, width=0.2, linetype=0) +
    geom_point(size=2.5)+
  coord_flip() +
  geom_hline(yintercept = 0)+
  xlab("")+
  ylab("Estimate") +
  scale_y_continuous(limits=c(-0.75, 0.3))
  #facet_wrap(~method) 

  #########CogZ
  colours <- c( "#339900",  "#FF9326","#D92121")
p3 <- dat.plot.base.clean %>% 
  filter(np_group=='PD') %>% 
    filter(method=='BIANCA') %>% 
  ggplot(aes(x=(wmh+1),y=global_z)) +
  #geom_violin()+
  geom_point(aes(color=diagnosis), alpha=0.6) +
  scale_colour_manual (values=colours,name = "diagnosis") + 
  scale_x_log10('1 + White Matter Hyperintensities (ml)',limits=c(0.9, 100)) +
  ylab("Global cognitive ability (z score)") +
  #facet_wrap(~method,nrow=1) +
  stat_smooth(aes(x=(wmh+1),y=resid),method="lm", color='black') +
  theme(legend.position="none")


#####UPDRS 3
p4 <-  dat.base.pd.updrs %>% 
  filter(np_group=='PD') %>% 
    filter(method=='BIANCA') %>% 
  ggplot(aes(x=(wmh+1),y=Part_III)) +
  #geom_violin()+
  geom_point(aes(color=diagnosis), alpha=0.6) +
  scale_colour_manual (values=colours,name = "diagnosis") + 
  scale_x_log10('1 + White Matter Hyperintensities (ml)',limits=c(0.9, 100)) +
  ylab("Motor impairment (UPDRS-III)") +
  #facet_wrap(~method,nrow=1) +
  stat_smooth(aes(x=(wmh+1),y=resid),method="lm", color='black') +
  theme(legend.position="none")



#####Arrange the grid


ggarrange(p2,p2.5,p3,p4, ncol=2,nrow=2,labels="AUTO")

ggsave("BIANCA_Figure.pdf",width=10,height=8) 

```


